Aspects of Dealing with Imperfect Data in Temporal Databases

### 2012-03-14 trunk/book/sources/temporalDatabases.tex
update in the bibliography, and in the graph for the example in the fuzzy time part. Update also in the commercial database systems, the addition of secondo as a temporal database system.

%
% Temporal databases
%
A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained. In subsections \ref{subsubsec:primary-key} and \ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed. Finally, subsection \ref{Comm-temp} presents an overview of some commercial temporal database systems.

@@

\emph{Valid times} are usually supplied by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}. Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support \cite{Dyreson1994}.

@@

\subsection{\label{subsubsec:primary-key}Primary Keys in Valid-time Relation Design}
Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is ususally chosen as primary key. The values of a tuple for these attributes will then uniquely determine this tuple and no two distinct tuples may have the exact same values for every attribute in this primary key. Next to attributes unrelated to time, a valid-time relation schema will typically contain one or more attributes which model the valid-time aspects and behavior of the real objects and concepts modelled by the relation schema. In this work, these attributes are called \emph{valid-time attributes}. In valid-time relation extentions, distinct tuples can exist containing the exact same values for every attribute except the valid-time attributes. These distinct tuples represent distinct versions of the same real object or concept, valid during different time periods. To allow the existence of such tuples when designing a valid-time relation using a relational database model, the most common solution is to include the valid-time attributes in the primary key.
%One of the first tasks in the design of a database is the choice for the primary key. In a temporal database, the most common solution for the choice of the primary key is to extend the primary key with the temporal value. Nevertheless, when dealing with uncertainty in the temporal values, the solution is usually to create a version identifier otherwise the primary key have uncertainty which should be avoided.

@@

\subsection{\label{subsubsec:consistency}Consistency in Valid-time Relation Querying}
TODO\\
TODO\\
TODO\\

The solution presented in subsection~\ref{subsubsec:primary-key} concerns relation design and consists of including the valid-time attributes in the primary key. Unfortunately, implementing this solution as such allows for the existence of records whose values imply inconsistencies with respect to the modelling of reality.

@@

The inconsistency problem can now be solved by replacing this statement with the next:

@@

The resulting relation is visualized in table~\ref{table:example-database-update}.

@@

\subsection{\label{Comm-temp}Commercial Temporal Database Systems}
Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} presents some information on some of the more well-known temporal DBMS and provides references for more information.

Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in OracleDB. On another note, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported). Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining. The most complete implementation is Workspace Manager.

@@

\begin{table}
\centering
\caption{Commercial Temporal Database Systems. }
\begin{tabular}{c c c c c c }
\hline
\textbf{Name} & \textbf{Time Supported} & \textbf{Comments} & \textbf{Reference}  \\ \hline
Oracle Workspace Manager & VT and TT. & Package for Oracle DB. & \cite{oracle2009}\\
TimeDB & VT and TT. & Interface for Oracle DB. & \cite{timedb2005}\\
Postgree Temporal & VT. & Package for Postgree SQL. & \cite{posgree2009}\\
Teradata & VT and TT. & Used for data-mining. & \cite{teradata2011}\\
\hline 
\end{tabular}
\label{table:commercial-temporal-db}

>>>>

%
% Temporal databases
%
A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained. In subsections \ref{subsubsec:primary-key} and \ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed. Finally, subsection \ref{Comm-temp} presents an overview of some commercial temporal database systems.

@@

\emph{Valid times} are usually provided by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}. Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support \cite{Dyreson1994}.

@@

\subsection{\label{subsubsec:primary-key}Primary Keys in Valid-time Relation Design}
Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is ususally chosen as primary key. The values of a tuple for these attributes will then uniquely determine this tuple, hence no two distinct tuples may have the exact same values for every attribute in this primary key. Next to attributes unrelated to time, a valid-time relation schema will typically contain one or more attributes which model the valid-time aspects and behavior of the real objects and concepts modelled by the relation schema. In this work, these attributes are called \emph{valid-time attributes}. In valid-time relation extentions, distinct tuples can exist containing the exact same values for every attribute except the valid-time attributes. These distinct tuples represent distinct versions of the same real object or concept, valid during different time periods. To allow the existence of such tuples when designing a valid-time relation using a relational database model, the most common solution is to include the valid-time attributes in the primary key.
%One of the first tasks in the design of a database is the choice for the primary key. In a temporal database, the most common solution for the choice of the primary key is to extend the primary key with the temporal value. Nevertheless, when dealing with uncertainty in the temporal values, the solution is usually to create a version identifier otherwise the primary key have uncertainty which should be avoided.

@@

\subsection{\label{subsubsec:consistency}Consistency in Valid-time Relation Content Modification}
The solution presented in subsection~\ref{subsubsec:primary-key} concerns relation design and consists of including the valid-time attributes in the primary key. Unfortunately, implementing this solution as such allows for the existence of records whose values imply inconsistencies with respect to the modelling of reality.

@@

\noindent
The inconsistency problem can now be solved by replacing this statement with:

@@

\noindent
The resulting relation is visualized in table~\ref{table:example-database-update}.

@@

\subsection{\label{Comm-temp}Commercial Temporal Database Systems}
Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} gives an overview of some of the more well-known temporal DBMS and provides references for more information.

Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in OracleDB. On another note, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported). Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining. Secondo~\cite{Dieker2000} is an extensible database system in which the core of the database may be replaced by a customized algebras. It is designed for non-standard applications and it support both valid and transaction times.

 The most complete implementation is Workspace Manager.

@@

\begin{table}
\centering
\caption{Commercial Temporal Database Systems. }
\begin{tabular}{c c c c c c }
\hline
\textbf{Name} & \textbf{Time Supported} & \textbf{Comments} & \textbf{Reference}  \\ \hline
Oracle Workspace Manager & VT and TT. & Package for Oracle DB. & \cite{oracle2009}\\
TimeDB & VT and TT. & Interface for Oracle DB. & \cite{timedb2005}\\
Postgree Temporal & VT. & Package for Postgree SQL. & \cite{posgree2009}\\
Teradata & VT and TT. & Used for data-mining. & \cite{teradata2011}\\
Secondo & VT and TT. & Spatio-temporal database. & \cite{Guting} \\
\hline 
\end{tabular}
\label{table:commercial-temporal-db}
